---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  echo = TRUE
)
```

# SeedMaker

<!-- badges: start -->
[![R-CMD-check](https://github.com/hdhshowalter/SeedMaker/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hdhshowalter/SeedMaker/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/hdhshowalter/SeedMaker/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/hdhshowalter/SeedMaker/actions/workflows/test-coverage.yaml)
[![lint](https://github.com/hdhshowalter/SeedMaker/actions/workflows/lint.yaml/badge.svg)](https://github.com/hdhshowalter/SeedMaker/actions/workflows/lint.yaml)
<!-- badges: end -->

{SeedMaker} is a mechanism for easily generating and organizing a collection of seeds from a single seed, which may be subsequently used to ensure reproducibility in processes/pipelines that utilize multiple random components (e.g., trial simulation).

## Installation

Install the released version of {SeedMaker} from CRAN:

``` r
install.packages("SeedMaker")
```

Or install the development version of {SeedMaker} from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("hdhshowalter/SeedMaker")
```

## Examples

The following calls are representative of the most common ways to engage `seed_maker()`:

``` {r}
library(SeedMaker)

#simple call
seed_maker(
  seed = 1234,
  level_names = "sim",
  n_per_level = 5
)

#slightly more complex call
seed_maker(
  seed = 1234,
  level_names = c("sim", "replicate"),
  n_per_level = c(5, 10)
)

#slightly more complex call with non-indexed names at level 2
seed_maker(
  seed = 1234,
  level_names = list("sim", c("component_a", "component_b", "component_c")),
  n_per_level = c(5, 3)
)

#simple call that generates seed values between 1 and 100 instead of between
#1 and 1000000
seed_maker(
  seed = 1234,
  level_names = "sim",
  n_per_level = 5,
  sample_vector = list(1:100)
)

#simple call that returns named vector instead of list
seed_maker(
  seed = 1234,
  level_names = "sim",
  n_per_level = 5,
  flatten = TRUE
)
```
